Model-based variable selection for regression with highly correlated variables.

### 2014-06-11 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@637 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Numerical results on simulated datasets} \label{sectionsimul}

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{reshatY} and \ref{resrobust} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods respectively for general case and extreme cases. Figures gives both mean and inter-quartile intervals of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$ and those of the sub-regressions are successively set to $(0.1,0.3,0.5,0.7,0.99)$. Variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$. The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign. $\forall j \in I_r, p_1^j=2$ (sub-regressions of length 2) and we have $p_r=16$ sub-regressions. The datasets were then scaled so that covariates $X_r$ don't have a greater variance or mean.
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$. %(see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
	We compare different methods with and without CorReg as a pre-treatment. All the results are provided by the CorReg package. In the following figures, the thickness represent various values of $n$, the thicker the greater, in $(30,50,100,400)$ so we start from $n<p$.
		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure?}
			The first criterion is $\psi(\boldsymbol{X},S)$ which is maximized in the MCMC. But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found $\hat{S}$:
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_r\cap \hat{I}_r|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_r|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_r|-TL$
				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple.
				%\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
			\end{itemize}

>>>>

\section{Numerical results on simulated datasets} \label{sectionsimul}
	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$. (see figure \ref{toutOLS}).
	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{reshatY} and \ref{resrobust} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods respectively for general case and extreme cases. Figures gives both mean and inter-quartile intervals of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$ and those of the sub-regressions are successively set to $(0.1,0.3,0.5,0.7,0.99)$. Variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$. The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign. $\forall j \in I_r, p_1^j=2$ (sub-regressions of length 2) and we have $p_r=16$ sub-regressions. The datasets were then scaled so that covariates $X_r$ don't have a greater variance or mean.
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.

	We compare different methods with and without CorReg as a pre-treatment. All the results are provided by the CorReg package. In the following figures, the thickness represent various values of $n$, the thicker the greater, in $(30,50,100,400)$ so we start from $n<p$.
		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure?}
			The first criterion is $\psi(\boldsymbol{X},S)$ which is maximized in the MCMC. But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found $\hat{S}$:
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_r\cap \hat{I}_r|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_r|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_r|-TL$
				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple.
				%\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
			\end{itemize}
